---
title: "SI2020_clustering_exercise_v2"
author: "PFR"
date: "August 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## --------------------------------------------------
##    PFR data prep exercises for clustering
##    using UCI Data Repository winedata
## --------------------------------------------------

```{r}
#use setwd("c:/your-path-to-your-project/data/")

X = read.table('winedata.csv',
                      header=TRUE,sep=",",
                      stringsAsFactors = TRUE)  #try TRUE
Y = X[,'Class']
X = subset(X,select=-c(Class))

plot(X[,'Alcohol'],X[,'Flavas'])
```


## Normalize data

```{r get numeric columns, echo=TRUE}

X_mncntr = scale(X, center=TRUE, scale= TRUE) #normalize to z-scores
head(X_mncntr,2)
```


## Run Kmeans for several K values


```{r kmeans, echo=TRUE}

# Run kmeans for 20 values of K
kg=matrix(0,20,2)
for (i in 1:20){
  ktest=kmeans(X_mncntr,i,20,5);
  kg[i,1]=i;
  kg[i,2]=ktest$tot.withinss;  #total withn SSE 
  }
plot(kg[,1],kg[,2],main='kmeans within cluster SS')

#try
#str(ktest)

```

## Now, let's plot the cluster mean values

```{r try 4 clusters, echo=TRUE}
#get Kmeans for 4 clusters 
k4_results     = kmeans(X_mncntr,4,20,5)

#get color scheme
col2use        = c('red','blue','black','yellow')

#get cluster assignment in colors
colassignments = col2use[k4_results$cluster]

plot(1:dim(X_mncntr)[2],k4_results$center[1,],type='l',col=col2use[1])

for (i in seq(2,4)){
  points(1:dim(X_mncntr)[2],k4_results$center[i,],type='l',col=col2use[i])
  }

```

## Now, let's combine the Kmeans cluster information with SVD 

## Exercise: rerun with different components - what does it say about the data and predictability?

```{r combine Kmeans cluster with SVD info, echo=TRUE}
#get SVD of X_mncntr
Xsvd = svd(X)

X_proj = as.matrix(X_mncntr) %*% Xsvd$v[,c(1,2)] 
#or just use Xsvd$u[,c(1,2)] 
#or use different components beside 1 and 2

plot(X_proj[,1],X_proj[,2],col=colassignments,main='data pts project to 1,2 SVD components, colored by kmeans')

# to plot center points, first project them into components
c3 = k4_results$centers%*% Xsvd$v[,1:3] 
points(c3[,1],c3[,2],pch=8,cex=2)   
#Try different components?


```

## Now try coloring by class (using Y mean centered)


```{r plot with Y color}

#Y was created in SVD exercise, use it to select 2 colors

#get class assignment in colors
colassignments = col2use[Y]

plot(X_proj[,1],X_proj[,2],col=colassignments,main='data pts project to 1,2 SVD components, colored by class')




```
## Extra:  Run a classification model, and color the above scatter plots by each class and correct predictions, or each class and incorrect predictions - what do you think you'll see

